---
title: 'STA302H1S / 1001 HS Autumn 2019 Assignment # 1'
author: "Xinrui Wang"
Student Number: '1003830063'
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
Group Code: ' I25'
---

Q1a

1)Equation: $\text{average season rating}_i $ = $d_{1990}*\mu_{1990}+d_{2000}*\mu_{2000}+d_{2010}*\mu_{2010}+\epsilon_i$
where di's are dummy variables for decade i
2) Assumption for ANOVA:
1. Errors are independent (observations are independent) 
2. Errors are normally distributed with E[$\epsilon_i$] = 0
3. Constant variance (homoscedasticity), var[$\epsilon_i$] = $\sigma^2$

Q1b
$H_0$: $\mu_{1990} = \mu_{2000} = \mu_{2010}$, the average season rating for crime shows is the same decade to decade
$H_a$: at least one of the average season ratings for crime shows is different from the others

Q1c
```{r, echo=TRUE, eval=TRUE}
library(tidyverse)

# load crimeshow data 
crime_show_data <- readRDS("crime_show_ratings.RDS")
# Side by side box plots 
crime_show_data %>% ggplot(aes(x = decade, y = season_rating)) + geom_boxplot() + ggtitle("Boxplots of average rating by decade for crime TV shows")

# Facetted histograms 
crime_show_data %>% ggplot(aes(x = season_rating)) + geom_histogram(bins=20) + facet_wrap(~decade) + ggtitle("Histograms of average rating by decade for crime TV shows")
```

1)I prefer the boxplot since it is more straightforward: mean and the values between 25th and 75th percentile can be easily seen(shown by the lines and the box) and compared across decades

2)I would improve it by investigating the outliers. If they are anomalous, I would take them out and create a boxplot without outliers. 

3) No. The difference is small as average ratings of 3 decades are all around 8 from the boxplot.

Q1d
```{r}
attach(crime_show_data)

a <- aov(season_rating ~ decade, data = crime_show_data)

summary(a)
```
From the ANOVA, the p-value under the null hypothesis test is 0.238 which is larger than 0.05. This does not provide evidence against $H_0$. Therefore, the sample supports $H_0$: average season ratings for crime shows are the same decade to decade.

Q1e
```{r}
plot(a, 1)
plot(a, 2)
```
The first plot shows the residuals of the model against the fitted values and the second plot shows the standardized residuals against a theoretical normal distribution of residuals. 

From the first plot: no random pattern is seen and points are not centered around 0 (E($\epsilon_i$),indicated by the red line). Thus, the assumption of E($\epsilon_i$) = 0 is violated.

From the second plot: a staright line can be fitted into the points/the points clearly form a straight line, thus, the assumption of $\epsilon_i$ follows a normal distribution is satisfied.

```{r}
var = crime_show_data %>% 
  group_by(decade) %>% 
  summarise(var_rating = sd(season_rating)^2)
var

max(var[,2])/min(var[,2]) < 3
```
Here, $\frac{S_{max}^2}{S_{min}^2 }$ < 3. Therefore, the constant variance assumption of $\epsilon_i$ is satisfied.

1f
```{r}
l = lm(season_rating ~ decade)
summary(l)

# to get mean season rating of each decade
c(lm(season_rating ~ 1, filter(crime_show_data, decade == 1990))$coef, 
  lm(season_rating ~ 1, filter(crime_show_data, decade == 2000))$coef, 
  lm(season_rating ~ 1, filter(crime_show_data, decade == 2010))$coef)
```
The intercepts are exactly the mean seanson rating for each decade. For 1990s, mean = intercept, 7.9222. For the next decade, mean = intercept + 0.1368. Then, the next decade, mean = intercept + 0.1938.
That is, $\hat\mu_{1990s}$ = 7.922151, $\hat\mu_{2000s}$ = 8.058942, $\hat\mu_{2010s}$ = 8.115985.