---
title: "A1"
author: "Xinrui Wang"
date: "2020/1/18"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1a

1)Equation: $\text{average season rating}_i $ = $d_{1990}*\mu_{1990}+d_{2000}*\mu_{2000}+d_{2010}*\mu_{2010}+\epsilon_i$
where di's are dummy variables for decade i
2) Assumption for ANOVA:
1. Errors are independent (observations are independent) 
2. Errors are normally distributed with E[$\epsilon_i$] = 0
3. Constant variance (homoscedasticity), var[$\epsilon_i$] = $\sigma^2$

Q1b
$H_0$: $\mu_{1990} = \mu_{2000} = \mu_{2010}$, the average season rating for crime shows is the same decade to decade
$H_a$: at least one of the average season ratings for crime shows is different from the others

Q1c
```{r, echo=TRUE}
library(tidyverse)

# load crimeshow data 
crime_show_data <- readRDS("crime_show_ratings.RDS")
# Side by side box plots 
crime_show_data %>% ggplot(aes(x = decade, y = season_rating)) + geom_boxplot() + ggtitle("Boxplots of average rating by decade for crime TV shows")

# Facetted histograms 
crime_show_data %>% ggplot(aes(x = season_rating)) + geom_histogram(bins=20) + facet_wrap(~decade) + ggtitle("Histograms of average rating by decade for crime TV shows")
```

1)I prefer the boxplot since it is more straightforward: mean and the values between 25th and 75th percentile can be easily seen(shown by the lines and the box) and compared across decades

2)I would improve it by investigating the outliers. If they are anomalous, I would take them out and create a boxplot without outliers. 

3) No. The difference is small as average ratings of 3 decades are all around 8 from the boxplot.

Q1d
```{r}
attach(crime_show_data)

a <- aov(season_rating ~ decade, data = crime_show_data)

summary(a)
```
From the ANOVA, the p-value under the null hypothesis test is 0.238 which is larger than 0.05. This does not provide evidence against $H_0$. Therefore, the sample supports $H_0$: average season ratings for crime shows are the same decade to decade.

Q1e
```{r}
plot(a, 1)
plot(a, 2)
```
The first plot shows the residuals of the model against the fitted values and the second plot shows the standardized residuals against a theoretical normal distribution of residuals. 

From the first plot: no random pattern is seen and points are not centered around 0 (E($\epsilon_i$),indicated by the red line). Thus, the assumption of E($\epsilon_i$) = 0 is violated.

From the second plot: a staright line can be fitted into the points/the points clearly form a straight line, thus, the assumption of $\epsilon_i$ follows a normal distribution is satisfied.

```{r}
var = crime_show_data %>% 
  group_by(decade) %>% 
  summarise(var_rating = sd(season_rating)^2)
var

max(var[,2])/min(var[,2]) < 3
```
Here, $\frac{S_{max}^2}{S_{min}^2 }$ < 3. Therefore, the constant variance assumption of $\epsilon_i$ is satisfied.

1f
```{r}
l = lm(season_rating ~ decade)
summary(l)

# to get mean season rating of each decade
c(lm(season_rating ~ 1, filter(crime_show_data, decade == 1990))$coef, 
  lm(season_rating ~ 1, filter(crime_show_data, decade == 2000))$coef, 
  lm(season_rating ~ 1, filter(crime_show_data, decade == 2010))$coef)
```
The intercepts are exactly the mean seanson rating for each decade. For 1990s, mean = intercept, 7.9222. For the next decade, mean = intercept + 0.1368. Then, the next decade, mean = intercept + 0.1938.
That is, $\hat\mu_{1990s}$ = 7.922151, $\hat\mu_{2000s}$ = 8.058942, $\hat\mu_{2010s}$ = 8.115985.


```{r}
smokeFile = 'smokeDownload.RData' 
if(!file.exists(smokeFile)){ download.file('http://pbrown.ca/teaching/303/data/smoke.RData', smokeFile)} 
(load(smokeFile))
smokeFormats[
smokeFormats[,'colName'] == 'chewing_tobacco_snuff_or', c('colName','label')]
smokeSub = smoke[which(smoke$Age > 10 & !is.na(smoke$Race)), ]

smokeSub$ageC = smokeSub$Age - 16
smokeModel = glm(chewing_tobacco_snuff_or ~ ageC + RuralUrban + Race + Sex, data=smokeSub, family=binomial(link='logit'))
knitr::kable(summary(smokeModel)$coef, digits=3)
```

```{r}
logOddsMat = cbind(est=smokeModel$coef, confint(smokeModel, level=0.99)) 
oddsMat = exp(logOddsMat) 
oddsMat[1,] = oddsMat[1,] / (1+oddsMat[1,]) 

rownames(oddsMat)[1] = 'Baseline prob' 
knitr::kable(oddsMat, digits=3)
```
Q2a

1) In 'smokeModel', the response variable£¬$Y_i$, follows a binomial distribution. i.e. $Y_i$~Binomial(N,$\mu_i$), where N equals the totoal number of observations,20,716 and $\mu_i$ equals to the probability of  have used chewing tobacco, snuff, or dip on 1 or more days in the past 30 days,given $X_i$.

The model is fitted into a glm, precisely, a logistic model. Such that $log(\frac{\mu_i}{1-\mu_i})$ = $\boldsymbol{X_i\beta}$.

2) See the table below for definitions of variable $Y_i's$ and $X_{ij}'s$ where i = 1, ..., 20,716(total number of observations in 'smokeSub') and j = 1, ..., 8(total number of $\beta's$ other than $\beta_0$):

| Variable | Definition                                                                                                           |
|----------|----------------------------------------------------------------------------------------------------------------------|
| $Y_i$    | the number of people in the data that have used chewing tobacco, snuff, or dip on 1 or more days in the past 30 days |
| $X_{i1}$ | the age of person i - 16                                                                                             |
| $X_{i2}$ | a binary variable, equals to 1 when person i lives in the rural area; 0 if in the urban area                         |
| $X_{i3}$ | a binary variable, equals to 1 when the race of person i is 'Black'; 0 if not 'Black'                                |
| $X_{i4}$ | a binary variable, equals to 1 when the race of person i is 'Hispanic'; 0 if not 'Hispanic'                          |
| $X_{i5}$ | a binary variable, equals to 1 when the race of person i is 'Asian'; 0 if not 'Asian'                                |
| $X_{i6}$ | a binary variable, equals to 1 when the race of person i is 'Native'; 0 if not 'Native'                              |
| $X_{i7}$ | a binary variable, equals to 1 when the race of person i is 'Pacific'; 0 if not 'Pacific'                            |
| $X_{i8}$ | a binary variable, equals to 1 when the person i is female; 0 if male                                                |

Q2b
'Baseline prob' is the probability of 16-year-old, urban white men that used chewing tobacco, snuff, or dip on 1 or more days in the past 30 days.

Q2c

```{r,echo=TRUE,eval=FALSE}
newData = data.frame(Sex = rep(c('M','F'), c(3,2)), Race = c('white','white','hispanic','black','asian'), ageC = 0, RuralUrban = rep(c('Rural','Urban'), c(1,4)))

smokePred = as.data.frame(predict(smokeModel, newData, se.fit=TRUE, type='link'))[,1:2]

smokePred$lower = smokePred$fit - 3*smokePred$se.fit 
smokePred$upper = smokePred$fit + 3*smokePred$se.fit 

expSmokePred = exp(smokePred[,c('fit','lower','upper')])

knitr::kable(cbind(newData[,-3],100*expSmokePred/(1+expSmokePred)), digits=1)
```
1) See from the table above that rural white males have the highest probability of using chewing tobacco.(By comparing the fitted probability, upper and lower bonds of the 99% CI. Rural white males is the only group that has probability over 10%.) And see that the estimated/fitted value is inside the 99% CI, there is no evidence to reject the hypothesis. Therefore, our data support the hypothesis that rural white males are the group most likely to use chewing tobacco.

2) See the last 2 rows of the table, notice that the fitted probabilities of urban black and urban asian females (ethnic-minority urban women) are both inside the 99% CI. This provides evidence that data support the claim of less than half of one percent(0.5%) of ethnic-minority urban women and girls chew tobacco.

Q3a
```{r}
fijiFile = 'fijiDownload.RData'
if(!file.exists(fijiFile)){ download.file(
'http://pbrown.ca/teaching/303/data/fiji.RData', fijiFile)
} 
(load(fijiFile))
```

```{r}

fijiSub = fiji[fiji$monthsSinceM > 0 & !is.na(fiji$literacy),] 

fijiSub$logYears = log(fijiSub$monthsSinceM/12) 

fijiSub$ageMarried = relevel(fijiSub$ageMarried, '15to18') 
fijiSub$urban = relevel(fijiSub$residence, 'rural')
fijiRes = glm( children ~ offset(logYears) + ageMarried + ethnicity + literacy + urban, family=poisson(link=log), data=fijiSub) 

logRateMat = cbind(est=fijiRes$coef, confint(fijiRes, level=0.99)) 
knitr::kable(cbind( summary(fijiRes)$coef, exp(logRateMat)), digits=3)

#2nd model
fijiSub$marriedEarly = fijiSub$ageMarried == '0to15' 

fijiRes2 = glm( children ~ offset(logYears) + marriedEarly + ethnicity + urban, family=poisson(link=log), data=fijiSub) 

logRateMat2 = cbind(est=fijiRes2$coef, confint(fijiRes2, level=0.99)) 
knitr::kable(cbind(summary(fijiRes2)$coef, exp(logRateMat2)), digits=3)
```
Q3a

1) In 'fijiRes', the response variable£¬$Y_i$, follows the Poisson distribution. i.e. $Y_i$~Poisson($\mu$).
$\mu$ = E(Y) = var(Y)
The model is fitted into a Poisson glm such that $log(\mu)$ = $\boldsymbol{X_i\beta}$.

2) See the table below for definitions of variable $Y_i's$ and $X_{ij}'s$ where i = 1, ..., 4,908(total number of observations in 'fijiSub') and j = 1, ..., 16(total number of $\beta's$ other than $\beta_0$):

| Variable  | Definition                                                                                        |
|-----------|---------------------------------------------------------------------------------------------------|
| $Y_i$     | the number of children the woman has after being first married                                    |
| $X_{i1}$  | a binary variable, equals to 1 if the woman i ages between 0 and 15 when first married; 0 if not  |
| $X_{i2}$  | a binary variable, equals to 1 if the woman i ages between 18 and 20 when first married; 0 if not |
| $X_{i3}$  | a binary variable, equals to 1 if the woman i ages between 20 and 22 when first married; 0 if not |
| $X_{i4}$  | a binary variable, equals to 1 if the woman i ages between 22 and 25 when first married; 0 if not |
| $X_{i5}$  | a binary variable, equals to 1 if the woman i ages between 25 and 30 when first married; 0 if not |
| $X_{i6}$  | a binary variable, equals to 1 if the woman i ages greater than 30 when first married; 0 if not   |
| $X_{i7}$  | a binary variable, equals to 1 if woman i is Indian; 0 if not                                     |
| $X_{i8}$  | a binary variable, equals to 1 if woman i is European; 0 if not                                   |
| $X_{i9}$  | a binary variable, equals to 1 if woman i is part European; 0 if not                              |
| $X_{i10}$ | a binary variable, equals to 1 if woman i is Pacific Islander; 0 if not                           |
| $X_{i11}$ | a binary variable, equals to 1 if woman i is Routman; 0 if not                                    |
| $X_{i12}$ | a binary variable, equals to 1 if woman i is Chinese; 0 if not                                    |
| $X_{i13}$ | a binary variable, equals to 1 if woman i is 'other'; 0 if not                                    |
| $X_{i14}$ | a binary variable, equals to 1 if woman i is illiterate; 0 if literate                            |
| $X_{i15}$ | a binary variable, equals to 1 if woman i lives in the capital city of Suva; 0 if not             |
| $X_{i16}$ | a binary variable, equals to 1 if woman i lives in other urban areas; 0 if not                    |

Q3b
```{r}
lmtest::lrtest(fijiRes2, fijiRes)
```

1) Yes. 'fijiRes2' is nested in 'fijiRes'. Because: 1. predictors of model fijiRes2: offset(logYears), ethnicity and urban all can be found in fijiRes. 2.the predictor 'marriedEarly' is a simplified version of the predictor 'ageMarried' OR 'marriedEarly' is created from 'ageMarried' so that 'marriedEarly' can also be found in the model 'fijiRes'.
And the command lmtest::lrtest(fijiRes2, fijiRes) runs a likelihood ratio test comparing the two. 
See that the p-value is greater than 0.01, therefore, we fail to reject $H_0$. So that 'fijiRes2' explains the data roughly as well as 'fijiRes'.

2) There are 2 constraints:
1. the value of $\beta_{literacy}$ is 0 in 'fijiRes2'.
2. the value of $\beta_{marriedEarly}$ in 'fijiRes2' is the increase in log($\mu$) for a woman married early than not early, while the value of the other predictors are the same whereas values of $\beta_{ageMarried}$for various age groups in 'fijiRes' are increases in log($\mu$) for a woman in the corresponding age group than a woman between age 15 and 18.

Q3c
I will fit the data into the model 'fijiRes2'.
$H_0$: $\beta_{literacy}$ = $\beta_{ageMarried > 15}$ = 0
$H_a$: one of $\beta_{literacy}$ and $\beta_{ageMarried > 15}$ ¡Ù 0

I looked at the p-values. It is found that none of the p-values is less than 0.01. Therefore, there is have no evidence to reject $H_0$. This means the data supports that there was no way for married women to influence their birth intervals.